---
title: "Statistics_1x_3x_5x"
date: "`r format(Sys.time(), '%d %b %Y')`"
author: "Taylor Lab"
output: 
  html_document:
    toc: true
    code_folding: "hide"
    df_print: kable
    knitr:
      fig.width: 8
      fig.height: 6
---

### Load necessary libraries and packages
```{r, warning = F, message = F}
library(pacman)
pacman::p_load(data.table, ggplot2, lubridate, stringr, ggpubr, dplyr, cowplot, readxl, scales, knitr, tidyr, patchwork)
```


```{r, warning = F, message = F}

################################################################################

save   <- TRUE
figure <- "Figure 4/4F/"

################################################################################

# PREPROCESSING
Input_Directory <- file.path("/Volumes/TAYLOR-LAB/Synthetic Myddosome Paper/6_Manuscript/Source files/", figure)
NAME_KEY        <- fread("~/Documents/Github/Analysis_Synthethic_Myddosome/FigureScripts/Figure 4 & S6/Figure_4_ELISA_CL_KEY.csv", header = T) 

# load functions
source("~/Documents/Github/Analysis_Synthethic_Myddosome/FigureScripts/functions.R")

# process ELISA
plate_data_raw <- ELISA_Fx(Input_Directory, Output_Directory)
plate_data     <- left_join(plate_data_raw, NAME_KEY)
plate_data     <- plate_data %>% filter(CELL_LINE != "NA") %>% unique()

# ensure correct column type assignment
plate_data$CONDITION <- as.factor(plate_data$CONDITION)

################################################################################
# subset and normalize the Dose Response Data

plotting_data <- plate_data %>% group_by(CELL_LINE, CL_NAME_ON_PLOT, STIM_DAY, CONDITION, STIM_CONCENTRATION, Date, Plate)

baseline <- plotting_data %>%
  group_by(Plate) %>%
  summarise(baseline_control_value = min(Concentration))

plotting_data <- left_join(plotting_data, baseline) %>%
  mutate(Concentration_REDUCED = case_when(!is.na(baseline_control_value) ~ Concentration - baseline_control_value, TRUE ~ Concentration))

# Baseline-corrected per cell line, per condition, per stimulation and retrieve the max values to normalize for day-to-day differences.
control_mean_per_day <- plotting_data %>%
  group_by(CELL_LINE, CL_NAME_ON_PLOT, CONDITION, STIM_CONCENTRATION, Date, STIM_DAY, Plate) %>%
  reframe(Concentration_REDUCED = Concentration_REDUCED) %>%
  ungroup() %>%
  group_by(STIM_DAY) %>%
  reframe(control_MEASUREMENT = max(Concentration_REDUCED))

# Normalize the baseline-corrected values by dividing each value by the max value per stimulation day.
plotting_data <- left_join(plotting_data, control_mean_per_day) %>%
  group_by(CELL_LINE, CL_NAME_ON_PLOT, CONDITION, STIM_CONCENTRATION, Date, STIM_DAY) %>%
  mutate(Concentration_NORMALIZED = case_when(Concentration_REDUCED / control_MEASUREMENT < 0 ~ 0, TRUE ~ Concentration_REDUCED / control_MEASUREMENT),
         Relative_Intensity_mean = mean(Concentration_NORMALIZED)) %>%
  ungroup()

# Calculate the normalized means per grouped variables
plotting_means <- plotting_data %>%
  group_by(CELL_LINE, CL_NAME_ON_PLOT, 
           STIM_CONCENTRATION,
           CONDITION,
           Date, STIM_DAY, Plate,
           PLOTTING_COLOR) %>%
  distinct(Relative_Intensity_mean, STIM_DAY, .keep_all = TRUE) %>%
  ungroup()


# Perform a t-test to compare the means per cell line per concentration
concentration_values <- unique(plotting_means$STIM_CONCENTRATION)
stat.test_list       <- list()

for (concentration_val in concentration_values) {
  concentration_data <- plotting_means %>% filter(STIM_CONCENTRATION == concentration_val)
  stat_test_result   <- ggpubr::compare_means(Relative_Intensity_mean ~ CL_NAME_ON_PLOT, data = concentration_data, method = "t.test") %>% mutate(STIM_CONCENTRATION = concentration_val)
  stat.test_list[[as.character(concentration_val)]] <- stat_test_result
}
stat.test <- do.call(rbind, stat.test_list)

# Convert the STIM_CONCENTRATION column to numeric
stat.test$STIM_CONCENTRATION <- as.numeric(as.character(stat.test$STIM_CONCENTRATION))

# Order the dataframe based on the STIM_CONCENTRATION column
stat.test <- stat.test[order(-stat.test$STIM_CONCENTRATION), ]

# Create a data frame with unique CL_NAME_ON_PLOT and corresponding PLOTTING_COLOR
color_mapping <- unique(plotting_means[, c("CL_NAME_ON_PLOT", "PLOTTING_COLOR")])

# Merge the plotting_means data frame with the color_mapping data frame
plotting_means_merged <- plotting_means %>%
  left_join(color_mapping) %>%
  mutate(CL_NAME_ON_PLOT = factor(CL_NAME_ON_PLOT, levels = unique(plotting_means$CL_NAME_ON_PLOT)))

# Calculate the mean of means and standard error of the mean per cell line and per condition
plotting_stats <- plotting_data %>%
  group_by(CELL_LINE, CL_NAME_ON_PLOT, STIM_CONCENTRATION, PLOTTING_COLOR) %>%
  summarise(Relative_Intensity_sem  = sem(Relative_Intensity_mean),
            Relative_Intensity_mean = mean(Relative_Intensity_mean)) %>%
  distinct(.keep_all = TRUE)

################################################################################

# Deriving breaks from unique STIM_CONCENTRATION values
concentration_breaks <- sort(unique(plotting_stats$STIM_CONCENTRATION))
```


```{r, warning = F, message = F, fig.dim=c(12,7)}
# Plot
figure_4F_Dose_Response_ELISA <- 
  ggplot(data = plotting_stats, aes(x = STIM_CONCENTRATION, y = Relative_Intensity_mean, group = CL_NAME_ON_PLOT, fill = CL_NAME_ON_PLOT)) +
  geom_ribbon(aes(x = STIM_CONCENTRATION, 
                  ymin = Relative_Intensity_mean - Relative_Intensity_sem,
                  ymax = Relative_Intensity_mean + Relative_Intensity_sem,
                  fill = CL_NAME_ON_PLOT),
              alpha = 0.4, show.legend = F) +
  geom_path(size = 0.75, aes(col = CL_NAME_ON_PLOT)) +
  geom_text(data = subset(plotting_stats, STIM_CONCENTRATION == max(STIM_CONCENTRATION)), 
            aes(label    = gsub("CHARMS-sT6BM-", "", CL_NAME_ON_PLOT), 
                x        = max(STIM_CONCENTRATION), 
                y        = Relative_Intensity_mean,
                fontface = "bold"),
            color = setNames(unique(plotting_means_merged$PLOTTING_COLOR), unique(plotting_means_merged$CL_NAME_ON_PLOT)),
            hjust = -0.1, vjust = 0.5, size = 12) +
  labs(x = "IL-1 Conc. [ng/mL]", 
       y = "Relative Response") +
  scale_fill_manual(name  = "CL_NAME_ON_PLOT", values = setNames(unique(plotting_means_merged$PLOTTING_COLOR), unique(plotting_means_merged$CL_NAME_ON_PLOT))) +
  scale_color_manual(name = "CL_NAME_ON_PLOT", values = setNames(unique(plotting_means_merged$PLOTTING_COLOR), unique(plotting_means_merged$CL_NAME_ON_PLOT))) +
  guides(color = "none", fill = guide_legend(reverse = TRUE, ncol = 1)) +
  theme_cowplot(font_family = "Helvetica") + 
  theme(legend.position = c(0.05,0.9),
        legend.title = element_blank(),
        plot.margin = unit(c(20,20,20,20), 'mm')) +
  scale_x_log10(breaks = concentration_breaks, labels = concentration_breaks)

figure_4F_Dose_Response_ELISA
```


### Checking for Normal distrbution with QQ-Plots

QQ plot draws the correlation between a given data and the normal distribution.

```{r, echo = F, warning = F, message = F, fig.dim=c(8,2)}
pacman::p_load(tidyverse, ggpubr, rstatix)

# subset the means data
data <- plotting_means[, c("CL_NAME_ON_PLOT", "Relative_Intensity_mean", "STIM_CONCENTRATION", "PLOTTING_COLOR")]
# data$CL_NAME_ON_PLOT <- gsub("CHARMS-sT6BM-", "", x = data$CL_NAME_ON_PLOT)

data$CL_NAME_ON_PLOT <- as.factor(data$CL_NAME_ON_PLOT)

concentration_values <- unique(data$STIM_CONCENTRATION)

# Create a list to store ggplot objects
qq_plots <- list()

for (concentration_val in concentration_values) {
  concentration_data <- plotting_means %>% filter(STIM_CONCENTRATION == concentration_val)
  res_aov <- aov(Relative_Intensity_mean ~ CL_NAME_ON_PLOT, data = concentration_data)
  qq_plot <- ggqqplot(res_aov$residuals, id = FALSE) +
    ggtitle(concentration_val) +
    theme_cowplot(font_size = 10)
  
  qq_plots[[as.character(concentration_val)]] <- qq_plot
}

# Combine all plots into one using facet_wrap
combined_plot <- wrap_plots(qq_plots, ncol = length(concentration_values))

# Print the combined plot
print(combined_plot)

```

### Checking for Normal distrbution with a Shapiro-Wilk Test

Check normality assumption with a Shapiro-Wilk test for each group level. If the data is normally distributed, the p-value should be greater than 0.05.

```{r, warning = F, message = F}

shapiro.test_list <- list()

for (concentration_val in concentration_values) {
  
  concentration_data <- plotting_means %>% filter(STIM_CONCENTRATION == concentration_val)
  
  res_aov <- aov(Relative_Intensity_mean ~ CL_NAME_ON_PLOT, data = concentration_data)
  
  # do we have normally distributed values?
  # here we assume normal distribution, so p-values > 0.05 are GOOD!
  
  shapiro_test_result <- shapiro.test(res_aov$residuals)[["p.value"]]
  shapiro.test_list[[as.character(concentration_val)]] <- shapiro_test_result
}

shapiro.test <- as.table(unlist(shapiro.test_list)) %>%
  as.data.frame() %>%
  setnames(old = c("Var1", "Freq"), new = c("STIM_CONCENTRATION", "p")) %>%
  mutate(normally_distributed = case_when(p > 0.05 ~ T, T ~ F))

shapiro.test

```


### Checking for Homogneity of variance assumption with Levene

The Levene’s test will be used to check the homogeneity of variances.
p-value is > 0.05 indicate that there is no significant difference between variances across groups. Therefore, we assume the homogeneity of variances in the different concentration treatment groups.

```{r, warning = F, message = F}

levene.test_list <- list()

for (concentration_val in concentration_values) {
  
  concentration_data <- plotting_means %>% filter(STIM_CONCENTRATION == concentration_val)
  
  levene <- concentration_data %>% 
    levene_test(Relative_Intensity_mean ~ CL_NAME_ON_PLOT) %>%
    mutate(STIM_CONCENTRATION = concentration_val)
  
  levene.test_list[[as.character(concentration_val)]] <- levene$p
}

levene.test <- as.table(unlist(levene.test_list)) %>%
  as.data.frame() %>%
  setnames(old = c("Var1", "Freq"), new = c("STIM_CONCENTRATION", "p")) %>%
  mutate(equal_variances = case_when(p > 0.05 ~ T, T ~ F))

levene.test
```

All concentrations have p-values above 0.05, so there are no significant differences between groups.

### One-Way ANOVA & Tukey

#### ANOVA

Values show homogeneous variances and normal distributions, therefore a normal ANOVA test can be performed.

```{r, warning = F, message = F}

anova_test_results <- data %>%
  group_by(STIM_CONCENTRATION) %>%
  anova_test(Relative_Intensity_mean ~ CL_NAME_ON_PLOT)

# anova_test_results
subset(anova_test_results[,c("STIM_CONCENTRATION", "F", "ges", "p", "p<.05")], p < 0.05)

```

#### Tukey post-hoc

A significant one-way ANOVA is generally followed up by Tukey post-hoc tests to perform multiple pairwise comparisons between groups.

```{r, warning = F, message = F}

pwc_tukey_hsd <- data %>% 
  group_by(STIM_CONCENTRATION) %>%
  tukey_hsd(Relative_Intensity_mean ~ CL_NAME_ON_PLOT)

pwc_tukey_hsd$STIM_CONCENTRATION <- reorder(pwc_tukey_hsd$STIM_CONCENTRATION, as.numeric(pwc_tukey_hsd$STIM_CONCENTRATION))

# pwc_tukey_hsd[,c("STIM_CONCENTRATION", "p.adj", "p.adj.signif", "group1", "group2")] %>% arrange(desc(STIM_CONCENTRATION))

subset(pwc_tukey_hsd[,c("STIM_CONCENTRATION", "p.adj", "p.adj.signif", "group1", "group2")], p.adj < 0.05)  %>% arrange(desc(STIM_CONCENTRATION))

if (save) {
  save_to_data_tay <- file.path("/Volumes/TAYLOR-LAB/Synthetic Myddosome Paper/6_Manuscript/Source files/", figure)
  save_to_local    <- file.path("~/Desktop/Source files/", figure)
  save_to          <- ifelse(test = dir.exists(save_to_data_tay), yes = save_to_data_tay, no = save_to_local)
  
  if (dir.exists(save_to)) {print(paste0("Files will be saved to ", save_to))} else {dir.create(save_to, recursive = T); print(paste0("Files will be saved to ", save_to))}
  
  # save tables
  fwrite(pwc_tukey_hsd,  file.path(save_to, "significance_pwc_tukey.csv"))
}
```




